Unbalanced Mallows Models for Optimizing Expensive Black-Box Permutation Problems

### 2021-02-04 paper/main.tex
minor fix
update zenodo link
update css concepts

\newcommand{\supplement}{\href{http://doi.org/10.5281/zenodo.4167148}{doi:10.5281/zenodo.4167148}}

@@

% \author{Ekhiñe Irurozki\inst{1} \and Manuel López-Ibáñez\inst{2}\orcidID{0000-0001-9974-1295}}
% \institute{
%    Basque Center for Applied Mathematics\\
%    \email{eirurozki@bcamath.org}
%    \and
%    University of Málaga, Málaga, Spain\\
%    \email{manuel.lopez-ibanez@uma.es}
%  }
%    \renewcommand{\shortauthors}{B. Trovato et al.}
\begin{abstract}
%The abstract should briefly summarize the contents of the paper in
%150--250 words.
  Expensive black-box combinatorial optimization problems arise in practice
  when the objective function is evaluated by means of a simulator or a
  real-world experiment. Since each fitness evaluation is expensive in terms of
  time or resources, only a limited number of evaluations is possible,
  typically several orders of magnitude smaller than in non-expensive problems. In
  this scenario, classical optimization methods such as mixed-integer
  programming and local search are not useful.  In the continuous case,
  Bayesian optimization, in particular using Gaussian processes, has proven
  very effective under these conditions. Much less research is available in the
  combinatorial case. In this paper, we propose and analyze UMM, an
  estimation-of-distribution (EDA) algorithm based on a Mallows probabilistic
  model and unbalanced rank aggregation (uBorda). Experimental results on
  black-box versions of LOP and PFSP show that UMM is able to match, and
  sometimes surpass, the solutions obtained by CEGO, a Bayesian optimization
  algorithm for combinatorial
  optimization. %, specially when the budget of solutions evaluated approaches 400.
  Moreover, the computational complexity of UMM increases linearly with both
  the number of function evaluations and the permutation size.
%  an order of magnitude smaller than CEGO, which makes it specially useful when
  %both computation time and number of evaluations are of concern.
  \sloppy
  %\textcolor{red}{16 pages max, deadline: 1 November 2020}
\end{abstract}
%
% The code below should be generated by the tool at
% http://dl.acm.org/ccs.cfm
% Please copy and paste the code instead of the example below. 
%
\begin{CCSXML}
<ccs2012>
 <concept>
  <concept_id>10010520.10010553.10010562</concept_id>
  <concept_desc>Computer systems organization~Embedded systems</concept_desc>
  <concept_significance>500</concept_significance>
 </concept>
 <concept>
  <concept_id>10010520.10010575.10010755</concept_id>
  <concept_desc>Computer systems organization~Redundancy</concept_desc>
  <concept_significance>300</concept_significance>
 </concept>
 <concept>
  <concept_id>10010520.10010553.10010554</concept_id>
  <concept_desc>Computer systems organization~Robotics</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
 <concept>
  <concept_id>10003033.10003083.10003095</concept_id>
  <concept_desc>Networks~Network reliability</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
</ccs2012>  
\end{CCSXML}

\ccsdesc[500]{Computer systems organization~Embedded systems}
\ccsdesc[300]{Computer systems organization~Redundancy}
\ccsdesc{Computer systems organization~Robotics}
\ccsdesc[100]{Networks~Network reliability}

@@

The \emph{Permutation Flowshop Scheduling Problem (PFSP)} is defined by a matrix $[p_{ij}]$ of size $n \times M$ that gives the processing time of a job $i \in [n]$ on a machine $j \in [M]$. All jobs must be processed by all machines in the same order. Given a permutation of the jobs $\sigma$, $C_{i,j}$ denotes the completion time of job at position $\sigma(i)$ on machine $j$, and $C_{\max} = C_{n,M}$  is the completion time of the last job on the last machine, i.e., the makespan. The objective of the PFSP is to find:
\begin{equation}\label{eq:pfsp}
  \begin{split}
    \min_{\sigma \in S_n}\quad& f(\sigma) = C_{n,M}\\
    \text{s.t.}\quad & C_{1,j} = 0 \quad j\in [M], \qquad C_{i,1} = 0\quad i \in [n]\\
    &C_{i,j} = p_{\sigma(i),j} + \max\{C_{i-1, j}, C_{i, j-1}\}  \ \  i \in \{2,\dotsc,n\},\ j \in\{2,\dotsc, M\}\\
\end{split}
\end{equation}

>>>>

\newcommand{\supplement}{\href{http://doi.org/10.5281/zenodo.4500974}{doi:~10.5281/zenodo.4500974}}

@@

% \author{Ekhiñe Irurozki\inst{1} \and Manuel López-Ibáñez\inst{2}\orcidID{0000-0001-9974-1295}}
% \institute{
%    Basque Center for Applied Mathematics\\
%    \email{eirurozki@bcamath.org}
%    \and
%    University of Málaga, Málaga, Spain\\
%    \email{manuel.lopez-ibanez@uma.es}
%  }
%    \renewcommand{\shortauthors}{B. Trovato et al.}
\begin{abstract}
%The abstract should briefly summarize the contents of the paper in
%150--250 words.
  Expensive black-box combinatorial optimization problems arise in practice
  when the objective function is evaluated by means of a simulator or a
  real-world experiment. Since each fitness evaluation is expensive in terms of
  time or resources, only a limited number of evaluations is possible,
  typically several orders of magnitude smaller than in non-expensive problems. In
  this scenario, classical optimization methods such as mixed-integer
  programming and local search are not useful.  In the continuous case,
  Bayesian optimization, in particular using Gaussian processes, has proven
  very effective under these conditions. Much less research is available in the
  combinatorial case. In this paper, we propose and analyze UMM, an
  estimation-of-distribution (EDA) algorithm based on a Mallows probabilistic
  model and unbalanced rank aggregation (uBorda). Experimental results on
  black-box versions of LOP and PFSP show that UMM is able to match, and
  sometimes surpass, the solutions obtained by CEGO, a Bayesian optimization
  algorithm for combinatorial
  optimization. %, specially when the budget of solutions evaluated approaches 400.
  Moreover, the computational complexity of UMM increases linearly with both
  the number of function evaluations and the permutation size.
%  an order of magnitude smaller than CEGO, which makes it specially useful when
  %both computation time and number of evaluations are of concern.
  \sloppy
  %\textcolor{red}{16 pages max, deadline: 1 November 2020}
\end{abstract}
%
% The code below should be generated by the tool at
% http://dl.acm.org/ccs.cfm
% Please copy and paste the code instead of the example below. 
%
\begin{CCSXML}
<ccs2012>
<concept>
<concept_id>10002950.10003624.10003625.10003630</concept_id>
<concept_desc>Mathematics of computing~Combinatorial optimization</concept_desc>
<concept_significance>500</concept_significance>
</concept>
<concept>
<concept_id>10003752.10010061.10011795</concept_id>
<concept_desc>Theory of computation~Random search heuristics</concept_desc>
<concept_significance>500</concept_significance>
</concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Mathematics of computing~Combinatorial optimization}
\ccsdesc[500]{Theory of computation~Random search heuristics}

@@

The \emph{Permutation Flowshop Scheduling Problem (PFSP)} is defined by a matrix $[p_{ij}]$ of size $n \times M$ that gives the processing time of a job $i \in [n]$ on a machine $j \in [M]$. All jobs must be processed by all machines in the same order. Given a permutation of the jobs $\sigma$, $C_{i,j}$ denotes the completion time of job at position $\sigma(i)$ on machine $j$, and $C_{\max} = C_{n,M}$  is the completion time of the last job on the last machine, i.e., the makespan. The objective of the PFSP is to find:
\begin{equation}\label{eq:pfsp}
  \begin{split}
    \min_{\sigma \in S_n}\quad& f(\sigma) = C_{n,M}\\
    \text{s.t.}\quad & C_{1,j} = 0 \quad j\in [M], \qquad C_{i,1} = 0\quad i \in [n]\\
    &C_{i,j} = p_{\sigma(i),j} + \max\{C_{i-1, j}, C_{i, j-1}\}  \ \  i \in \{2,\dotsc,n\},\\
    &\hspace{14.5em} j \in\{2,\dotsc, M\}
  \end{split}
\end{equation}
